---
title: "all_alaska_soils"
author: "Weronika Konwent"
date: "2024-05-09"
output: html_document
---

## BNZ Soils
The purpose of this code is to collect age, site, and deciduous data for sites surveyed through the BNZ that ALREADY HAVE CN soil and age data. There is no actual soil data here, though. These sites are not necessarily RSN sites.

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Library/CloudStorage/OneDrive-NorthernArizonaUniversity/R/BNZ_RSN_WG")

library(ggplot2)
library(tidyverse)
library(dplyr)

```

Some of betsy's data in central Alaska from the FISL project. Much of Betsy's work was in deciduous stands, which make these sites good candidates for filling in soil gaps for young deciduous sites.
```{r betsy's data}
#read in betsy's soil that I received from Xanthe
black_soil <- read_csv("Data/raw/Black_soil/xandata043024.csv")

names(black_soil)
head(black_soil)

#use Xanthe's code to calculate deciduous index
black_soil %>%
  mutate(density.dec=density.as+density.br+density.bp, 
         density.con=density.bs+density.ws+density.unk.con, 
         rel.dens.con=density.con/(density.con+density.dec), 
         rel.bio.con=biomass.con/(biomass.con+biomass.dec), 
         rel.dens.dec=1-rel.dens.con, 
         rel.bio.dec=1-rel.bio.con,
         d.index=(rel.dens.dec+rel.bio.dec)/2, 
         class=ifelse(d.index<0.33, "spruce", 
                      ifelse(d.index>0.66, "decid", "mixed"))) -> black_soil.a

black_soil.a=as.data.frame(black_soil.a)

black_soil.b=select(black_soil.a, study, fire_scar, site, postfire.age, d.index, class)

#assign trajectory
black_soil.b %>%
  group_by(fire_scar, site, postfire.age, study)%>%
  summarize(d.index = mean(d.index, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(class=ifelse(d.index<0.33, "spruce", 
                      ifelse(d.index>0.66, "decid", "mixed"))) %>% 
  rename(sitename = fire_scar) -> black_soil.c

black_soil.c %>% 
  mutate(sitename = paste0(sitename, "_", site)) %>% 
  select(study, sitename, postfire.age, d.index, class) -> black_soil.c

```
Heather's sites are mostly intermediate deciduous. These data are also shared informaly, and are not archived. 
```{r Heather Alexander data}
ha_soil <- read_csv("Data/raw/Black_soil/HA_data/HAlexander_SOILS.csv")
ha_sites <- read_csv("Data/raw/Black_soil/HA_data/HA_treecats_forCpools_alltrees.csv")

names(black_soil)
head(black_soil)

#only keep sites for which we have soil data
ha_soil <- ha_soil[complete.cases(ha_soil$CN), ]

#let's clean up ha_site a bit
ha_sites %>%
  mutate(study = "HA_chrono",
         subplot = gsub("chrono_", "", subplot),
         plot = gsub("chrono_", "", plot)) %>% 
  rename(postfire.age = "StdAge") %>% 
  select(study, plot, subplot, DCM.cat, postfire.age) -> ha_sites

#we have to standardize age across sites
#jeremy wrote this code. yikes.
num_sites <- unique(ha_sites$plot)

age_vector <- vector()
for (i in 1:length(num_sites)) {
  age_vector <- c(age_vector,max(subset(ha_sites, plot == num_sites[i])[["postfire.age"]], na.rm = TRUE))
}
age_vector[is.infinite(age_vector)]<-NA
age_vector<-data.frame(Sites=num_sites,Age=age_vector)

for(i in 1:dim(ha_sites)[[1]]){
  ha_sites$postfire.age[i] <- subset(age_vector, Sites == ha_sites$plot[i])[['Age']]
}
rm(age_vector,num_sites,i)

#let's make a uniqueID that matches ha_site so that we can check if all of the ha_sites were sampled for soil
ha_soil <- unite(ha_soil, "plot", Site, Transect, sep = "")
ha_soil <- unite(ha_soil, "subplot", plot, Subsection, sep = "_")

#this way we get an observation for every soil sample
### could i just skip using ha_soil and only use ha_site?
ha_soils <- full_join(ha_soil, ha_sites, by = "subplot")
  
ha_soils %>% 
  rename( sitename = "plot") %>% 
  select(study, sitename, DCM.cat, postfire.age) %>% 
  group_by(study, sitename, DCM.cat, postfire.age) %>% 
  summarise(count = n(), .groups = 'drop') %>% 
  pivot_wider(names_from = DCM.cat, values_from = count, values_fill = list(count = 0)) %>% 
  ungroup() -> ha_soils

#remove NA row
ha_soils <- ha_soils[complete.cases(ha_soils$sitename), ]
ha_soils %>% 
  select(study, sitename, postfire.age, conifer, deciduous, mixed) -> ha_soils

#let's make a deciduous index out of count
ha_soils %>% 
  rowwise() %>%
  mutate(total_count = sum(c_across(conifer:mixed), na.rm = TRUE)) %>%
  ungroup() %>% 
  mutate(d.index = (conifer/total_count)*0.25 + (mixed/total_count)*0.5 + (deciduous/total_count)*0.75,
  class=ifelse(d.index<0.33, "spruce", 
                      ifelse(d.index>0.66, "decid", "mixed"))) -> ha_soils

#here we're randomly assigning an intermediate age to some of Heather's sites that don't have ages.   
#postfire.age = ifelse(is.na(postfire.age), 35, postfire.age)) -> test

```
These data are pulled from the BNZ archive, under Melanie Jean's work. Many sites are also Heather Alexander sites.
```{r new version of HA data}
#these data are pulled from the BNZ database
ha_jean <- read_csv("Data/raw/Black_soil/801_Jean_SiteEnvironmentalData.csv")

#let's calculate deciduous index from biomass and density
ha_jean %>%
  mutate(density.dec=Pbdensity+Bpdensity+Ptdensity, 
         density.con=Pmdensity, 
         biomass.dec = Bpbiomass + Pb_biomass, Ptbiomass,
         biomass.con = Pmbiomass,
         rel.dens.con=density.con/(density.con+density.dec), 
         rel.bio.con=biomass.con/(biomass.con+biomass.dec), 
         rel.dens.dec=1-rel.dens.con, 
         rel.bio.dec=1-rel.bio.con,
         d.index=(rel.dens.dec+rel.bio.dec)/2, 
         class=ifelse(d.index<0.33, "spruce", 
                      ifelse(d.index>0.66, "decid", "mixed"))) -> ha_jean.a

ha_jean.a=as.data.frame(ha_jean.a)

#clean up the df to just include what we're interested in. note that we will want to recalculate the age so that it is accurate for 2024
ha_jean.b=select(ha_jean.a, SiteCode, BurnYear, d.index, class)

#let's fix the age here
current_year <- as.integer(format(Sys.Date(), "%Y"))
ha_jean.b %>%
  mutate(postfire.age = current_year - BurnYear) -> ha_jean.b

#ok, now let's combine this to the HA data so that we know which sites we have soil data for
values_jean <- ha_jean.b %>% distinct(SiteCode)
values_ha <- ha_soils %>% distinct(sitename)
common_ha <- intersect(values_jean$SiteCode, values_ha$sitename)

#filter ha_jean.b to keep only rows with site in common_ha
ha_jean.b %>%
  filter(SiteCode %in% common_ha) -> ha_jean.c

#ok, this is bonked, but what if I make an ha_chrono version with the d.index to complement the jean data
values_jean <- ha_jean.c %>% distinct(SiteCode)
values_ha <- ha_soils %>% distinct(sitename)
common_ha <- intersect(values_jean$SiteCode, values_ha$sitename)

ha_soils %>%
  filter(!(sitename %in% common_ha)) -> ha_soils.a
#ok, so actually turns out that all of these overflow sites don't have attached ages....so I think I'm going to ignore them.
#let's use ha_jean.c to represent HA data in the total dataset

#let's make ha_jean.c match our other dataset
ha_jean.c %>% 
  mutate(study = "ha_jean") %>% 
  rename(sitename = "SiteCode") %>% 
  select(study, sitename, postfire.age, d.index, class) -> ha_jean.c

```
So, we will use only ha_jean to indicate both HA and JEAN sites. 

We can also pull some TKN sites that have soil AND age data. Teresa's stuff is classically spruce, but we'll be particularly interested in her old spruce sites.
```{r TKN data}
#these data are also from the BNZ database
tkn_soil <- read_csv("Data/raw/321_Soil_CandN_2001.csv")
tkn_age <- read_csv("Data/raw/139_treecores_rings.csv")

names(tkn_soil)
head(tkn_soil)

tkn_soil %>%
  group_by(Site) %>%
  summarize(sample = n()) %>%
  mutate(study = "TKN") %>% 
  select(study, Site) -> tkn_soil

#find common unique_ids between the two datasets
common_age <- intersect(tkn_age$Site, tkn_soil$Site)
#we have 36 sites in age that we know have soil data

#let's keep just these 36 sites
tkn_age <- tkn_age[tkn_age$Site %in% common_age, ]

#add tkn age
tkn_age %>% 
  rename(tree_num = "Tree #",
         est_age = "Est. inner date of est.") %>% 
  select(Site, tree_num, est_age) -> tkn_age

tkn_age_site <- tkn_age %>%
  pivot_wider(names_from = tree_num, values_from = est_age, names_prefix = "tree")
#hmmmm, looks like there are some sites that have two entries for tree #4, or have a tree 4 and are missing a tree 3
#we need to do some cleaning up of these data

#what if we do a combination group by site and number
tkn_age_unique <- tkn_age %>%
  group_by(Site, tree_num) %>%
  mutate(row = row_number()) %>%
  ungroup()

#then we pivot, so that each site that has a duplicate tree number gets an extra row. 
tkn_age_site <- tkn_age_unique %>%
  pivot_wider(names_from = tree_num, values_from = est_age, names_prefix = "tree")

#then we can target each row 2 to change it to tree5
#this doesn't fix our problem with missing tree 3s, but that doesn't matter since we'll be taking summary stats anyway
tkn_age_site <- tkn_age_site %>%
  group_by(Site) %>%
  mutate(tree5 = ifelse(any(row == 2), lead(tree4), tree5)) %>%
  filter(row == 1) %>%
  ungroup()

tkn_age_site <- select(tkn_age_site, Site, tree1, tree2, tree3, tree4, tree5, tree6)
#now we can create summary stats
#we will again use out current year so that the ages are correct as this dataset ages

#let's make a current year value so that our ages are always current
current_year <- as.integer(format(Sys.Date(), "%Y"))
#the cores in tree_age were all collected in 2011

#average stand age
tkn_age_site$age_ave <- rowMeans(tkn_age_site[2:7], na.rm = TRUE)
tkn_age_site$age_ave <- current_year - tkn_age_site$age_ave
tkn_age_site$age_ave <- round(tkn_age_site$age_ave, 0)

tkn_age <- select(tkn_age_site, Site, age_ave)

tkn_age %>%
  mutate(study = "TKN",
         class = "spruce",
         d.index = "0") %>% 
  rename(postfire.age = age_ave,
         sitename = Site) -> tkn_age

# ### d index
# tree_inv <- read_csv("Data/working/rsn_bio_dens_bysite.csv")
# tree_inv <- rename(tree_inv, sitename = SITE)
# 
# #need to dd prelimIDs to tree_inv
# site_names_df <- read_csv("Data/working/RSN_site_names.csv")
# tree_inv <- full_join(tree_inv, site_names_df, by = "sitename")
# 
# #first let's filter out just tkn_age sites
# common_age <- intersect(tkn_age$sitename, tree_inv$prelimID)
# site_names_df <- site_names_df[site_names_df$sitename %in% common_age, ]
# tree_inv <- tree_inv[tree_inv$prelimID %in% common_age, ]
# 
# #let's keep site, d.index, decid.bio, decid.dens, and class
# decid_index <- select(tree_inv, sitename, prelimID, D.Index, Class)
# 
# decid_index %>% 
#   #rename columns for clarity
#   rename(d.index = D.Index,
#          class = Class,
#          site = sitename,
#          sitename = prelimID) %>% 
#   #make class data lower case
#   mutate(class = tolower(class)) -> decid_index
# 
# #now let's join this to tkn_age
# tkn_age <- full_join(tkn_age, decid_index, by = "sitename")
# tkn_age %>%
#   mutate(d.index = coalesce(as.character(d.index.y), as.character(d.index.x)),
#          class = coalesce(class.x, class.y)) -> tkn_age
# 
# tkn_age <- select(tkn_age, study, sitename, postfire.age, class, d.index)
# 
# #for use in making figures, let's keep age, decid.bio, decid.dens, D.Index and class
# bio_dens_age_bysite <- select(rsn_master, sitename, age, decid.bio, decid.dens, d.index, class)

```
An issue with Teresa's data is that we don't know what age represents. Age of last disturbance? Actual age? 

We can also pull some jfsp sites  that have soil data courtesy of leslie boby and anastasia pulak.
```{r Pulak / Boby data}
#read in file that anastasia made that combines data from her sampling of jfsp sites and leslie boby sampling of jfsp sites
#this file was shared informally by anastasia - it is not archived anywhere.
jfsp_soil <- read_csv("Data/working/JFSP.sample.inventory.csv")

names(jfsp_soil)
head(jfsp_soil)

jfsp_soil %>%
  #make study column
  mutate(study = "JFSP",
         traj.20 = tolower(traj.20),
         traj.20 = case_when(traj.20 == "deciduous" ~ "decid", TRUE ~ traj.20)) %>% 
  #rename columns for good joining
  rename(class = "traj.20",
         d.index = "decid.index",
         sitename = "site") -> jfsp_soil

#separate boby and ap sampling so that we can make ages for them
boby_soil <- select(jfsp_soil, study, sitename, d.index, class, sampled_2005)
ap_soil <- select(jfsp_soil, study, sitename, d.index, class, sampled_2023)

#keep only sites ap sampled in 2023, give them a postfire age
ap_soil %>% 
  filter(sampled_2023 != "no") %>% 
  mutate(postfire.age = 2023 - 2004) %>% 
  select(study, sitename, postfire.age, d.index, class) -> ap_soil

#give boby sites a postfire age
boby_soil %>% 
  mutate(postfire.age = 2005 - 2004) %>% 
  select(study, sitename, postfire.age, d.index, class) -> boby_soil

#join the two sampling years to one jfsp dataset
jfsp_soil <- rbind(ap_soil, boby_soil)

#i'm going to remove all "open" trajectories for now
jfsp_soil %>% 
  filter(class != "open",
         #4 sites do now have d.indexes. why?
         d.index != "NA") -> jfsp_soil

###what do we want to do about "open" trajectories
```

Combine all 4-5 disparate data sets.
```{r combining all data}

all_soils <- rbind(black_soil.c, tkn_age, jfsp_soil, ha_jean.c)
all_soils$d.index <- as.numeric(as.character(all_soils$d.index))

```

Figure comparing age and d.index. all of these sites also have CN data
```{r making figures}

#figure for d index vs stand age
f <- ggplot(all_soils, aes(x = postfire.age, y = d.index, color = class)) +
  geom_point() +
  theme_bw() +
  labs(x = "Stand Age", y = "Deciduous Index") +
  geom_point(size = 3, aes(fill = class, shape = study)) +
  scale_color_manual(values = c("gold1", "#77D153FF", "#443983FF", "#ADD8E6")) +
  scale_x_continuous(breaks = seq(0, 250, by = 10),
                     limits = c(0, 250),
                     labels = function(x) ifelse(x %% 50 == 0, as.character(x), "")) +
  scale_y_continuous(breaks = c(0, 0.25, 0.5, 0.75, 1))
f

```

```{r figure for sites we want to collect soils at}
#figure from above with potential cafi_rsn sites added
#we want to keep only those between 35 and 100 years old
bnz_cafi <- read_csv("Data/working/bnz_cafi.csv")
bnz_cafi <- subset(bnz_cafi, age_ave >= 35 & age_ave <= 100)

#to make nice shapes
study_shapes <- c("FiSL" = 17, "ha_jean" = 16, "JFSP" = 15, "TKN" = 18)
id_shapes <- c("CAFI" = 1, "RSN" = 2)

#to make nice legend
all_soils$study <- factor(all_soils$study, levels = c("FiSL", "ha_jean", "JFSP", "TKN"))
bnz_cafi$ID <- factor(bnz_cafi$ID, levels = c("CAFI", "RSN"))

#figure for D.index v age for sites that have CN data, with overlay of potential RSN and CAFI sites that do not have CN data
ggplot() +
  geom_point(data = all_soils, aes(x = postfire.age, y = d.index, color = class, shape = study), size = 3) +
  geom_point(data = bnz_cafi, aes(x = age_ave, y = d.index, color = class, shape = ID), size = 3, fill = NA) +
  theme_bw() +
  labs(x = "Stand Age", y = "Deciduous Index") +
  scale_color_manual(values = c("gold1", "#77D153FF", "#443983FF")) +
  scale_shape_manual(values = c(study_shapes, id_shapes)) +
  scale_x_continuous(breaks = seq(0, 250, by = 10),
                     limits = c(0, 250),
                     labels = function(x) ifelse(x %% 50 == 0, as.character(x), "")) +
  scale_y_continuous(breaks = c(0, 0.25, 0.5, 0.75, 1)) +
  guides(color = guide_legend(order = 1), shape = guide_legend(order = 2)) -> f.1
f.1

```
Notice here that the graphs are capped at 250 years on the x axis. we are missing some of the oldest teresa sites, because of this. 
At this point, in October 2024, there was associated CN data for 1 RSN site (UP4C) that we know of. This data came from Guelph, completeing loss on ignition (LOI) method on a permafrost core 
