---
title: "BNZ RSN inventory"
output: html_document
date: "2023-10-12"
editor_options: 
  chunk_output_type: console
---

#### This was originally made by Xanthe to create deciduous indexess for RSN sites. This file is a duplicate of Xanthe's that I have been modifying to make figures. My code begins at line 240.
```{r setup, include=FALSE}
rm(list=ls())

library(dplyr)

bnz <- read_csv("Data/working/BNZMasterSeries.csv")

```

#Determine how many sites in each year
###2018 was the last year that most sites (91 out of 114) were inventoried
### Sites were inventoried in 19 different years from 1989 to 2023

```{r organizing}
names(bnz)
head(bnz)
str(bnz)

bnz$SPECIES <- toupper(bnz$SPECIES) 
bnz$SITE <- toupper(bnz$SITE) 
levels(as.factor(bnz$SPECIES))
bnz$SPECIES <- gsub(" ", "", bnz$SPECIES)
str(bnz)

bnz$DBH=as.numeric(bnz$DBH)
bnz$DBH.new <- replace(bnz$DBH, bnz$DBH<0, NA)
bnz$DBH.new=as.numeric(bnz$DBH.new)

bnz.sites=bnz%>%
  group_by(YEAR) %>%
  summarize(unique_types = n_distinct(SITE))

n_distinct(bnz$SITE)
n_distinct(bnz$YEAR)


```


## calculate species specific density and basal in 2018
```{r bnz}
########live trees in 2018
bnz.2018=bnz%>%
  subset(YEAR==2018 & DA==1)


########live trees in 2013
bnz.2013=bnz%>%
  subset(YEAR==2013&DA==1)

######### merge 2018 and 2013 - keeping all 2018 and only unique from 2013
library(data.table)
bnz.new=unique(rbindlist(list(bnz.2018, bnz.2013)), by = c("SITE", "PLOT", "TREE", "SPECIES"))

head(bnz.new)

bnz.new=mutate(bnz.new, BA.tree = 3.142*(DBH.new/2)^2)

str(bnz.new)
bnz.new$SPECIES=as.factor(bnz.new$SPECIES)
levels(bnz.new$SPECIES)


bnz.new=bnz.new%>%
  mutate(bio.tree=ifelse(SPECIES=="POPTRE", 134.1*(DBH.new)^2.26, 
                        ifelse(SPECIES=="POPBAL", 133.71*(DBH.new)^2.29,  
                               ifelse(SPECIES=="BETNEO", 164.18*(DBH.new)^2.29,  
                                      ifelse(SPECIES=="PICMAR", 271.46*(DBH.new)^1.84, 
                                             ifelse(SPECIES=="PICGLA", 96.77*(DBH.new)^2.4,
                                                    ifelse(SPECIES=="PICEA"|SPECIES=="LARLAR", ((271.46*(DBH.new)^1.84)*0.5) + ((96.77*(DBH.new)^2.4)*0.5), NA)))))))
                                      
head(bnz.new)       


########make sure it worked
try=bnz.new%>%
  group_by(YEAR)%>%
  summarize(count=n())

#remove trees that have NA for DBH.new, so that the NAs don't throw off our calculations
bnz.new <- bnz.new[complete.cases(bnz.new$DBH.new), ]


#############calculate plot density
bnz.a=bnz.new%>%
  group_by(SITE, PLOT, SPECIES)%>%
  summarize(tree.num=n_distinct(TREE), 
            density=tree.num/100, 
            basal.area.tree=sum(BA.tree),
            ba=basal.area.tree/100, 
            biomass.area=sum(bio.tree),
            bio=biomass.area/100)

levels(as.factor(bnz.a$SPECIES))

max(bnz.a$ba, na.rm=TRUE)
min(bnz.a$ba, na.rm=TRUE)
max(bnz.a$density, na.rm=TRUE)
min(bnz.a$density, na.rm=TRUE)
max(bnz.a$bio, na.rm=TRUE)
min(bnz.a$bio, na.rm=TRUE)

library(ggplot2)
library(ggforce)
head(bnz.a)



#######get rid of species that are not species
bnz.c=subset(bnz.a, SPECIES=="BETNEO"|SPECIES=="LARLAR"|SPECIES=="PICEA"|SPECIES=="PICGLA"|SPECIES=="PICMAR"|SPECIES=="POPBAL"|SPECIES=="POPTRE")

#change sites to a factor
levels(as.factor(bnz.c$SITE)) 

#### plot density by sites and species - one page for 12 sites
g=ggplot(bnz.c, aes(x=SPECIES, y=density, fill=SPECIES))+facet_wrap_paginate(~ SITE, scales="free",ncol = 4, nrow = 4, page = 1)
g=g+theme_bw()+geom_boxplot()
g

g=ggplot(bnz.c, aes(x=SPECIES, y=density, fill=SPECIES))+facet_wrap_paginate(~ SITE, scales="free",ncol = 4, nrow = 4, page = 2)
g=g+theme_bw()+geom_boxplot()
g

g=ggplot(bnz.c, aes(x=SPECIES, y=density, fill=SPECIES))+facet_wrap_paginate(~ SITE, scales="free",ncol = 4, nrow = 4, page = 3)
g=g+theme_bw()+geom_boxplot()
g

g=ggplot(bnz.c, aes(x=SPECIES, y=density, fill=SPECIES))+facet_wrap_paginate(~ SITE, scales="free",ncol = 4, nrow = 4, page = 4)
g=g+theme_bw()+geom_boxplot()
g

g=ggplot(bnz.c, aes(x=SPECIES, y=density, fill=SPECIES))+facet_wrap_paginate(~ SITE, scales="free",ncol = 4, nrow = 4, page = 5)
g=g+theme_bw()+geom_boxplot()
g

g=ggplot(bnz.c, aes(x=SPECIES, y=density, fill=SPECIES))+facet_wrap_paginate(~ SITE, scales="free",ncol = 4, nrow = 4, page = 6)
g=g+theme_bw()+geom_boxplot()
g

#####################
#can do the same thing with basal area

g=ggplot(bnz.c, aes(x=SPECIES, y=ba, fill=SPECIES))+facet_wrap_paginate(~ SITE, scales="free", ncol = 4, nrow = 4, page = 1)
g=g+theme_bw()+geom_boxplot()
g

g=ggplot(bnz.c, aes(x=SPECIES, y=ba, fill=SPECIES))+facet_wrap_paginate(~ SITE, scales="free", ncol = 4, nrow = 4, page = 2)
g=g+theme_bw()+geom_boxplot()
g

g=ggplot(bnz.c, aes(x=SPECIES, y=ba, fill=SPECIES))+facet_wrap_paginate(~ SITE, scales="free", ncol = 4, nrow = 4, page = 3)
g=g+theme_bw()+geom_boxplot()
g

g=ggplot(bnz.c, aes(x=SPECIES, y=ba, fill=SPECIES))+facet_wrap_paginate(~ SITE, scales="free", ncol = 4, nrow = 4, page = 4)
g=g+theme_bw()+geom_boxplot()
g

g=ggplot(bnz.c, aes(x=SPECIES, y=ba, fill=SPECIES))+facet_wrap_paginate(~ SITE, scales="free", ncol = 4, nrow = 4, page = 5)
g=g+theme_bw()+geom_boxplot()
g

g=ggplot(bnz.c, aes(x=SPECIES, y=ba, fill=SPECIES))+facet_wrap_paginate(~ SITE, scales="free", ncol = 4, nrow = 4, page = 6)
g=g+theme_bw()+geom_boxplot()
g

#########

#####################
#can do the same thing with biomass

g=ggplot(bnz.c, aes(x=SPECIES, y=bio, fill=SPECIES))+facet_wrap_paginate(~ SITE, scales="free", ncol = 4, nrow = 4, page = 1)
g=g+theme_bw()+geom_boxplot()
g
```

```{r bnz}

library(tidyr)
names(bnz.c)
head(bnz.c)

bnz.b =bnz.c %>%  
select(SITE,PLOT,SPECIES,tree.num, density, ba, bio)%>%  
    pivot_wider(names_from = SPECIES, values_from = c(tree.num, density, ba, bio))

names(bnz.b)

bnz.d=bnz.b%>%
  replace(is.na(.),0) %>% 
  mutate(decid.dens=density_POPTRE+density_BETNEO+density_POPBAL, 
         conif.dens=density_PICGLA+density_PICEA+density_PICMAR, 
         decid.ba=ba_POPTRE+ba_BETNEO+ba_POPBAL, 
         conif.ba=ba_PICGLA+ba_PICEA+ba_PICMAR, 
         decid.bio=bio_POPTRE+bio_BETNEO+bio_POPBAL, 
         conif.bio=bio_PICGLA+bio_PICEA+bio_PICMAR, 
         rel.conif.dens=conif.dens/(conif.dens+decid.dens), 
         rel.conif.bio=conif.bio/(conif.bio+decid.bio), 
         rel.decid.dens=1-rel.conif.dens, 
         rel.decid.bio=1-rel.conif.bio, 
         D.Index=(rel.decid.dens+rel.decid.bio)/2, 
         Class=ifelse(D.Index<0.33, "Spruce", 
                      ifelse(D.Index>0.66, "Decid", "Mixed")))

f=ggplot(bnz.d, aes(x=conif.dens, y=decid.dens, color=Class))+theme_bw()
f= f+ labs(x="Coniferous Density (stems m-2)", y="Deciduous Density (stems m-2)")
f=f + geom_point(size=3,aes(fill=Class, shape=Class)) + scale_color_manual(values=c("gold1","#77D153FF", "#443983FF"))
f


f1=ggplot(bnz.d, aes(x=conif.ba, y=decid.ba, color=Class))+theme_bw()
f1= f1+ labs(x="Coniferous Basal Area", y="Deciduous Basal Area")+ geom_point(size=3,aes(fill=Class, shape=Class)) + scale_color_manual(values=c("gold1","#77D153FF", "#443983FF"))
f1

f2=ggplot(bnz.d, aes(x=conif.bio, y=decid.bio,color=Class))+geom_point()+theme_bw()
f2= f2+ labs(x="Coniferous Biomass (g m-2)", y="Deciduous Biomass (g m-2)")+ geom_point(size=3,aes(fill=Class, shape=Class)) + scale_color_manual(values=c("gold1","#77D153FF", "#443983FF"))
f2

names(bnz.d)
f3=ggplot(bnz.d, aes(x=rel.decid.bio, y=rel.decid.dens, color=Class))+geom_point()+theme_bw()
f3= f3+ labs(x="Relative Decid Biomass", y="Relative Deciduous Density")+ geom_point(size=3,aes(fill=Class, shape=Class)) + scale_color_manual(values=c("gold1","#77D153FF", "#443983FF"))
f3

library(ggpubr)
f4=ggarrange(f, f1, f2, f3, nrow=2, ncol=2,common.legend = TRUE )
f4
```

#### Summarize to site level instead of plot 
#### Combine this site level data with the tree age data!!
```{r bnz}

bnz.d=as.data.frame(bnz.d)

bnz.e=select(bnz.d, SITE, tree.num_PICMAR:D.Index)
bnz.e=bnz.e%>%
  group_by(SITE)%>%
  summarize_all(mean)

bnz.e=mutate(bnz.e, Class=ifelse(D.Index<0.33, "Spruce", 
                      ifelse(D.Index>0.66, "Decid", "Mixed")))

#save this is a working dataset as biomass and density on site level
write_csv(bnz.e, file = "Data/working/rsn_bio_dens_bysite.csv")
#then I do some changing of the code in my rsn_masterfile script
#and we read in the new data that includes age
bnz.f <- read_csv("Data/working/bio_dens_age_bysite.csv")
bnz.f$ID <- "RSN"

#remove pesky NA
bnz.f <- bnz.f[complete.cases(bnz.f$class), ]
bnz.f <- bnz.f[complete.cases(bnz.f$age_ave), ]

#change biomass to kg
bnz.f$decid.bio <- bnz.f$decid.bio/1000

#read in cafi data
cafi.b <- read_csv("Data/working/cafi.b.csv")
cafi.b %>% 
  mutate(ID = "CAFI",
         Class = tolower(Class)) %>% 
  rename(sitename = "site",
         age_ave = "StdAge", 
         d.index = "D.Index",
         class = "Class") %>% 
  select(ID, sitename, age_ave, d.index, class) -> cafi.c
cafi.c$sitename=as.character(cafi.c$sitename)

#cafi.d will be a version of cafi that only includes sites we may be interested in adding
cafi.c %>% 
  filter(age_ave > 54, d.index > 0.1) -> cafi.d

#let's make a quick df for just the two heather alexander sites we want to add
ha <- data.frame(
  sitename = c("ha_7", "ha_8"),
  latitude = c(64.902, 63.368),
  longitude = c(-148.275, -143.534),
  age_ave = c(87, 119),
  class = c("mixed", "aspen"),
  density.con = c(0.00, 0.28),
  density.dec = c(0.26, 0.24),
  biomass.con = c(0, 42),
  biomass.dec = c(19949, 12713))

ha %>%
  mutate(rel.dens.con=density.con/(density.con+density.dec), 
         rel.bio.con=biomass.con/(biomass.con+biomass.dec), 
         rel.dens.dec=1-rel.dens.con, 
         rel.bio.dec=1-rel.bio.con,
         d.index=(rel.dens.dec+rel.bio.dec)/2, 
         class=ifelse(d.index<0.33, "spruce", 
                      ifelse(d.index>0.66, "decid", "mixed"))) -> ha.a

ha.a %>% 
  select(sitename, latitude, longitude, age_ave, d.index, class) -> ha.a

#save ha.a for helena use
write_csv(ha.a, file = "Data/working/ha.a.csv")

#let's make a bnz.f that can be combined with cafi
bnz.f %>% 
  select(ID, sitename, age_ave, d.index, class) -> bnz.g
bnz.g$sitename=as.character(bnz.g$sitename)

#these are the rsn sites + potentialy cafi sites that are deciduous and over 54 yo.
bnz_cafi <- bind_rows(bnz.g, cafi.c)
#let's save it for use with making graphs
write_csv(bnz_cafi, file = "Data/working/bnz_cafi.csv")

#remove pesky NA
#bnz_cafi <- bnz_cafi[complete.cases(bnz_cafi$class), ]

#df <- bind_rows(cafi.b, bnz.f)
#df$ID_Class <- paste(df$ID, df$Class, sep = "_")

#cafi and rsn plot
f <- ggplot(bnz_cafi, aes(x = age_ave, y = d.index, color = class)) +
  geom_point() +
  theme_bw() +
  labs(x = "Stand Age", y = "Deciduous Index") +
  geom_point(size = 3, aes(fill = class, shape = ID)) +
  scale_color_manual(values = c("gold1", "#77D153FF", "#443983FF")) +
  scale_x_continuous(breaks = seq(0, 250, by = 10),
                     limits = c(0, 250),
                     labels = function(x) ifelse(x %% 50 == 0, as.character(x), ""))
f

#chosen CAFI sites over RSN
#for legend
shape_va
lues <- c(2, 16)
aspect_ratio <- 1

ggplot() +
  geom_point(data = bnz.f, aes(x = age_ave, y = d.index, color = class, shape = "current RSN"), alpha = 0.75, size = 3) +
  geom_point(data = cafi.d, aes(x = age_ave, y = d.index, color = class, shape = "potential RSN"), alpha = 0.75, size = 3, fill = NA) +
  geom_point(data = ha.a, aes(x = age_ave, y = d.index, color = class, shape = "potential RSN"), alpha = 0.75, size = 3, fill = NA) +
  theme_bw() +
  labs(x = "Stand Age", y = "Deciduous Index") +
  scale_color_manual(values = c("gold1", "#77D153FF", "#443983FF")) +
  scale_shape_manual(values = shape_values) +
  scale_x_continuous(breaks = seq(0, 2250, by = 10),
                     limits = c(0, 250),
                     labels = function(x) ifelse(x %% 50 == 0, as.character(x), "")) +
  scale_y_continuous(breaks = c(0, 0.25, 0.5, 0.75, 1)) +
  theme(legend.position = "bottom") + 
  guides(color = guide_legend(order = 1), 
         shape = guide_legend(title = "site status", 
                              labels = c("current RSN", "potential RSN"), 
                              order = 2)) +
  theme(aspect.ratio = 1) -> f
f

#D.Index v age
f <- ggplot(bnz.f, aes(x = age_ave, y = d.index, color = class)) +
  geom_point() +
  theme_bw() +
  labs(x = "Stand Age", y = "Deciduous Index") +
  geom_point(alpha = 0.5, size = 3, aes(fill = class, shape = class)) +
  scale_color_manual(values = c("gold1", "#77D153FF", "#443983FF", "#87CEEB")) +
  scale_x_continuous(breaks = seq(0, 250, by = 10),
                     limits = c(0, 250),
                     labels = function(x) ifelse(x %% 50 == 0, as.character(x), ""))
f

#decid.bio v age
f <- ggplot(bnz.f, aes(x = age_ave, y = decid.bio, color = Class)) +
  geom_point() +
  theme_bw() +
  labs(x = "Stand Age", y = "Deciduous Biomass (kg m-2)") +
  geom_point(size = 3, aes(fill = Class, shape = Class)) +
  scale_color_manual(values = c("gold1", "#77D153FF", "#443983FF")) +
  scale_x_continuous(breaks = seq(0, 250, by = 10),
                     limits = c(0, 250),
                     labels = function(x) ifelse(x %% 50 == 0, as.character(x), "")) +
  scale_y_continuous(breaks = seq(0, 16, by = 4),  # Set breaks every 2 units
                     limits = c(0, 16))  # Limiting the y-axis from 0 to 16
f

#decid.dens v age
f <- ggplot(bnz.f, aes(x = age_ave, y = decid.dens, color = Class)) +
  geom_point() +
  theme_bw() +
  labs(x = "Stand Age", y = "Deciduous Density") +
  geom_point(size = 3, aes(fill = Class, shape = Class)) +
  scale_color_manual(values = c("gold1", "#77D153FF", "#443983FF")) +
  scale_x_continuous(breaks = seq(0, 250, by = 10),  # Set breaks every 10 units
                     limits = c(0, 250),  # Set x-axis limits
                     expand = c(0, 0),  # Remove padding
                     labels = function(x) ifelse(x %% 50 == 0, as.character(x), ""))  + # Show numbers every 50 units
  scale_y_continuous(breaks = seq(0, 1, by = 0.25),  # Set breaks every 0.25 units
                     limits = c(0, 1))  # Set y-axis limits
f

```
